--- 
title: " Introduction to Probability <br> \\newline First Edition"
author: ""
# Include title and author for HTML/Gitbook output
site: bookdown::bookdown_site
output:
  bookdown::gitbook
documentclass: krantzNoCorner
classoption: krantz2
colorlinks: yes
fontsize: 12pt
bibliography: ["referece2025.bib"]
biblio-style: econPeriod
link-citations: yes
github-repo: https://github.com/openprobabilitytext/introprob
description: "Introduction to Probability is a freely available textbook for MATH350 students.
  -  This version contains variety of examples to learn the probability concepts in your own phase.
  -  The book can be downloaded as a pdf."
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = T, echo = TRUE, comment=NA, message = FALSE, warning = FALSE, fig.align = "center")
HtmlEval <- knitr:::is_html_output()
PdfEval  <- knitr:::is_latex_output()

```


```{r echo = FALSE}
#  Some Switches
QUIZ              <- TRUE
ANIMATION         <- TRUE
SHOW_PDF          <- TRUE
SHOW_EXAMPLE_SOLN <- TRUE
SHOW_EXER_SOLN    <- TRUE
ANIMATIONHOOK     <- 'gifski'
GLOSSARY          <- FALSE
DIST_TABLES_LATEX <- FALSE

if (PdfEval){
QUIZ              <- FALSE
ANIMATION         <- FALSE
SHOW_PDF          <- FALSE
SHOW_EXER_SOLN    <- FALSE
ANIMATIONHOOK     <- NULL
GLOSSARY          <- TRUE
DIST_TABLES_LATEX <- TRUE
numAnimation = 1
}
# Switches that control the number of animations in Chapter 6
# Use these for pdf output - maybe later
#numAnimation = 1
#numAnimation2 = 2
#numAnimation3 = 2

roundCom  <- function(number, digits) {
  rounded_number <- round(number, digits)
  formatted_number <- format(rounded_number, big.mark = ",", nsmall = digits)
  return(formatted_number)
}

CodeFontSmall <- function(A){ '\\scalefont{0.75}'    }
CodeFontLarge <- function(A){ '\\scalefont{1.33333}' }

SolnBegin <- function(A){ if (HtmlEval) {                    knitr::asis_output(paste("*Example Solution.*")) } 
  else { '\\begin{graybox}\\scalefont{0.90}
    \\textbf{Example Solution.}' }  }
SolnEnd <- function(A){ 
   if (PdfEval) { '\\end{graybox}' } }

LObjBegin <- function(A){ if (PdfEval){ '\\begin{graybox}\\scalefont{0.90}' } else
  {knitr::asis_output(paste("***")) } }
LObjEnd <- function(A){  if (PdfEval) { '\\end{graybox}' } else
  {knitr::asis_output(paste("***")) } }

Refer <- function(A){
if (PdfEval) { 'hello' } } #Inside: You can learn more about the R code for this example at the online version of this book.

#Only return argument for glossary function
Gloss <- function(term, section){
  return(term)
}

HideExample <- function(arg1,arg2){
  return("")
}

HideRCode <- function(arg1,arg2){
  return("")
}

HideProofTheory <- function(arg1,arg2){
  return("")
}


```


# Preface {-}

<p style="text-align:right;">
*Date: `r format(Sys.Date(), "%d %B %Y")`*</p>

<!--chapter:end:index.Rmd-->

# Fundamentals of Probability and Its Axioms

## Combinations and Permutations
Counting plays a very important role in probability. In probability, we often deal with sets, and counting methods such as combinations and permutations help us find the number of elements in a set. Specifically, in probability, we deal with a set called the **sample space** which is the set of all possible outcomes of some random experiment and **events**, which are subsets of the sample space. Typically, the sample space is denoted by $S$ or $\Omega.$ Consider the following example:

**Example 1.1.**
Suppose a fair coin is flipped twice.What is the probability of flipping at least one head?


`r HideExample('Example 1.1.', 'Show Example Solution')`

`r SolnBegin()` The sample space of this experiment can expressed as $S=\{HH,HT,TH,TT\}$. Now, consider the event:
$$A=\{\text{At least one outcome is a head}\}$$
Notice that $A\subset S$ ($A$ is a subset of $S$). Since $A=\{HH,HT,TH\},$ we can find the probability as follows:
$$P(A)= \frac{\#\text{elements in A}}{\#\text{elements in S}}=\frac{3}{4}.$$
`r SolnEnd()`

</div>

<!--chapter:end:02-counting.Rmd-->

# Basic Probability

<!--chapter:end:03-basics.Rmd-->



```{r echo = FALSE}
#  Default Widths for Latex if latexFont = 10
#HtmlEval <- FALSE
#PdfEval <- TRUE
ColWidth3  <- "2.4cm"
ColWidth4  <- "1.8cm"
ColWidth5  <- "1.6cm"
ColWidth55 <- "1.5cm"
ColWidth6  <- "1.4cm"
ColWidth7  <- "1.3cm"
ColWidth8  <- "1.1cm"
ColWidth9  <- "1.0cm"
ColWidth10 <- "0.9cm"
ColWidth11 <- "0.8cm"
ColWidth12 <- "0.7cm"
ColWidth14 <- "0.5cm"

#  Default Widths for Latex if latexFont = 8
#HtmlEval <- FALSE
#PdfEval <- TRUE
  if (HtmlEval){
 #  Default Widths for HTML
     # ColWidth3  <- "3.6cm"    
     # ColWidth4  <- "3.6cm"
     # ColWidth5  <- "3.2cm"
     # ColWidth55 <- "3.0cm"
     # ColWidth6  <- "2.8cm"
     # ColWidth7  <- "2.6cm"
     # ColWidth8  <- "2.2cm"
     # ColWidth9  <- "2.0cm"
     # ColWidth10 <- "1.8cm"
     # ColWidth11 <- "1.7cm"   
     # ColWidth12 <- "1.6cm"          
     # ColWidth14 <- "1.0cm" }
     ColWidth3  <- "3.6cm"    
     ColWidth4  <- "3.2cm"
     ColWidth5  <- "3.0cm"
     ColWidth55 <- "2.8cm"
     ColWidth6  <- "2.6cm"
     ColWidth7  <- "2.4cm"
     ColWidth8  <- "2.2cm"
     ColWidth9  <- "2.0cm"
     ColWidth10 <- "1.8cm"
     ColWidth11 <- "1.7cm"   
     ColWidth12 <- "1.6cm"          
     ColWidth14 <- "1.0cm" }

#  Function to Create a Style for Generating Tables
TableGen1 <-function(TableData,TextTitle,Align=`r`, Digits=0, ColumnSpec=1,
                     ColumnSpec0=1, ColWidth0= "2.5cm", 
                     BorderRight=1, ColWidth= "1.5cm", latexFont = 10,
                     htmlFont = 12, Border_Right = FALSE){
  if (HtmlEval){
     TextTitle1 <- paste0('**', TextTitle,'**', collapse='')
     if (is.na(TextTitle)){TextTitle1 <- NA}
     kableExtra::kbl(TableData, caption=TextTitle1, align = Align,
                     booktabs = T, digits=Digits) %>%
     kableExtra::kable_classic(full_width = F, font = htmlFont, 
                               html_font = "Cambria") %>%
     kable_styling(bootstrap_options = c("striped", "condensed"))  %>%
     kableExtra::column_spec(ColumnSpec0, width_min = ColWidth0) %>%
     kableExtra::column_spec(ColumnSpec,  width_min = ColWidth)    %>%
     kableExtra::column_spec(BorderRight, border_right = Border_Right)
   } else  
   { ColumnSpec <- ColumnSpec+1
     TextTitle1 <- paste0('\\textbf{',TextTitle,'}', collapse='')
     if (is.na(TextTitle)){TextTitle1 <- NA}
     kableExtra::kbl(TableData, caption = TextTitle1, align = Align,
                     booktabs = T, digits=Digits,escape = FALSE) %>%
    kableExtra::kable_styling(latex_options = c("striped", "condensed") ,
                              font = latexFont, protect_latex= TRUE)   %>%
     kableExtra::column_spec(ColumnSpec, width =  ColWidth)     %>%
    kableExtra::column_spec(BorderRight, border_right = Border_Right)
   }  
}
  
```




```{r echo = FALSE}
Gloss_data <- read.csv("GlossFct.csv", header = TRUE)
Gloss_data$Term = trimws(tolower(Gloss_data$Term))

A1 <- '<a href="#" class="tooltip" style="color:green">*'
A2 <- '*<span style="font-size:8pt">'
A3 <- '</span></a>'

Glossfunction <- function(term){
  temp2 <- which(Gloss_data == trimws(tolower(term)), arr.ind = TRUE)
  ifelse(length(temp2) == 0, return(''), return(Gloss_data[temp2[[1]],2]))
}

GlossfunctionBySection <- function(term, section){
  temp2 <- which(Gloss_data$Term == trimws(tolower(term)) & Gloss_data$Section == section)
  ifelse(length(temp2) == 0, return(''), return(Gloss_data[temp2,2]))
}

#Gloss <- function(term){
#  return(paste(A1,term,A2,Glossfunction(term),A3, sep=""))
#}

Gloss <- function(term, section){
  if (missing(section)) {
    section = FALSE
  }
  
  if (section == FALSE) {
    return(paste(A1,term,A2,Glossfunction(term),A3, sep=""))
  } else {
    return(paste(A1,term,A2,GlossfunctionBySection(term, section),A3, sep=""))
  }
}

```




```{r echo = FALSE}
B1 <- '<h5 style="text-align: center;"><a id="displayExample.'
B2 <- ' href="javascript:toggleEX(\'toggleExample.'
B3 <- '\',\'displayExample.'
B4a <- '\');"><i><strong>'
B4b <- '</strong></i></a> </h5><div id="toggleExample.'
B5 <- '" style="display: none">'

HideExample <- function(arg1,arg2){
  return(paste(B1,arg1,'"',B2,arg1,B3,arg1,B4a,arg2,B4b,arg1,B5, sep=""))
} 

C1 <- '<h5 style="text-align: center;"><a id="displayCode.'
C2 <- ' href="javascript:togglecode(\'toggleCode.'
C3 <- '\',\'displayCode.'
C4a<- '\');"><i><strong>'
C4b<- '</strong></i></a> </h5><div id="toggleCode.'
C5 <- '" style="display: none">'

HideRCode <- function(arg1,arg2){
  return(paste(C1,arg1,'"',C2,arg1,C3,arg1,C4a,arg2,C4b,arg1,C5, sep=""))
}

D1 <- '<h5 style="text-align: center;"><a id="displayTheory.'
D2 <- ' href="javascript:toggleTheory(\'toggleTheory.'
D3 <- '\',\'displayTheory.'
D4a<- '\');"><i><strong>'
D4b<- '</strong></i></a> </h5><div id="toggleTheory.'
D5 <- '" style="display: none">'

HideProofTheory <- function(arg1,arg2){
  return(paste(D1,arg1,'"',D2,arg1,C3,arg1,D4a,arg2,D4b,arg1,D5, sep=""))
}

```


```{r echo = FALSE, eval = PdfEval}
#Only return argument for glossary function
Gloss <- function(term, section){
  return(term)
}

HideExample <- function(arg1,arg2){
  return("")
}

HideRCode <- function(arg1,arg2){
  return("")
}

HideProofTheory <- function(arg1,arg2){
  return("")
}

```



<!--chapter:end:Glossary.Rmd-->


```{r include = FALSE}
library("actuar")
library("alabama")
library("epitools")
library("GB2")
library("pander")
library("VGAM")
library("psych")
library("copula")
library("statmod")

library("tseries")
library("boot")
library("scales")
library("rpart")
library("rpart.plot")
library("RColorBrewer")
library("forecast")
library("splines")
library("MASS")
library("survival")
library("Hmisc")
library("goftest")
library("gifski")
library("invgamma")
library("ChainLadder")
library("knitr")
library("kableExtra")
options(kableExtra.html.bsTable = FALSE)
library("magick")
library("pdftools")
library("dplyr")
library("GofCens")
library("Metrics")
library("webexercises")

library("invgamma")
library("patchwork")
library("mcmcse")
library('gganimate')
library('transformr')
#library('CASdatasets')
#library('insuranceData')
library('ggplot2')
library("nimble")
library("gridExtra")
library("grid")

# Install as needed

# install.packages("CASdatasets", repos = "http://cas.uqam.ca/pub/", type="source")


```

<!--chapter:end:Packages.Rmd-->


```{r echo = FALSE, eval = PdfEval}
QUIZ = FALSE
ANIMATION = FALSE 
SHOW_PDF = FALSE
ANIMATIONHOOK = NULL
GLOSSARY = TRUE
DIST_TABLES_LATEX = TRUE
# Switches that control the number of animations in Chapter 6
# Use these for pdf output - maybe later
#numAnimation = 1
#numAnimation2 = 2
#numAnimation3 = 2

#Only return argument for glossary function
Gloss <- function(term, section){
  return(term)
}

HideExample <- function(arg1,arg2){
  return("")
}

HideRCode <- function(arg1,arg2){
  return("")
}

HideProofTheory <- function(arg1,arg2){
  return("")
}

```


<!--chapter:end:PdfOutput.Rmd-->

